### 2020-02-05 rebuttal/response.txt
hm

Reviewer A

1. It is too bad that the experimental evaluation does not include some DPLL-based
string solvers.

Response: to the best of our knowledge, Z3 and CVC4 are the best DPLL-based string solver that are still well-maintained.

2. In Section 9, it would be nice to break down the X results further:
how many "unknown" vs timeouts vs incorrect results vs crashes? For
the latter two categories, did this work uncover new bugs in some of
the solvers? Please clarify.

Response: hlin

3. In Section 9, would using a timeout value different from 10 seconds
change the results significantly? This question is partly answered in
the text, but a scatter plot showing the solving time for each
instance and comparing Z3-PFA with other solvers would provide more
information.

Response: hlin

Reviewer B

1. The concept of flat automata for solving string constraints has already been proposed [1]. The solver framework based on both over- and under-approximation has also been proposed in [1]. I understand that this paper proposes a parametric version of flat automata (PFA). However, I do not see the key insight of the PFA. Why can't FA be applied in the problem domain since PFA is just FA plus recording the loop size/length?

Response: the difference between PFA and FA is that 
(1) PFA handle the transition labels symbolically and hence avoid the alphabet explosion problem. For example, if Unicode (>60000 symbols) is used, one PFA transition labeled by a variable might represent >60000 transitions of an FA. 
(2) In PFA, symbols are also natural numbers and hence can be used directly in the conversion between strings and integers. In FA, this requires multiple if-then-else statements, and these translate to predicates for which a DPLL engine must guess the truth values, which is costly.  
  As as simple experiment, we encoded the problem of finding an n-digit string whose integer value is larger than 42 using FA and also using PFA. 
For the case of n=10 digits, the DPLL solver needed to make 90 decisions when run on the integer formula created by the FA-based approach, while with the one from the PFA-approach, no decision at all needed to be made (the actual runtimes were insignificant since in the time limited time given, we only managed to handcode FA encodings of a relatively small formula. However, the difference in the complexity of the SMT queries seems obvious, and the FA formulae that would be generated from our actual benchmark would be much larger).

2. I can't entirely agree with the stated hardness of handling string-number conversion in solvers. In particular, lines 195-197 state that the challenge of dealing with string-number constraint is that the problem is undecidable. Well, undecidability is an over-used term in this context. Recall that the problem of solving string constraints itself is already very hard (see the undecidability discussion in section 2 of [1]). Therefore, the problem with string-number conversion, intuitively, is undecidable.

Response:

3. How significant is it to introduce the string-number conversion?

Response: JavaScript makes far more use of such conversions than other languages to which SMT solvers have traditionally been applied, such as C.  Every access to the state of any modeled object relies on it, since array indices get turned into strings.  Since JavaScript is pervasive, such checking is very significant.

4. With that said, the current paper does not give any constructive argument to discuss the hardness of the problem. I think perhaps a better way is to show that the popular solvers based on FA could not handle the string-number conversion. That gives a strong motivation for PFA as well as this work in general.
Response: FA-based approach can in theory be turned to support str-int conversion, but would be more expensive than using PFA. See the 2nd item of response 1. In practice, str-int conversion is not implemented in the existing FA-based tool Trau. (should we add more empirical support? saying that we tried FA+str-int conversion and PFA_str-int conversion is much faster?)

5. In the introduction, the description of the motivating example in Figure 2 is extremely confusing. When discussing the example for the first time (lines 285-300), I don't see why $A_x$ is projected in this way. In the example, we have only $y\in(12)^+$. How exactly do you infer that $A_x$ should have a similar structure?

Response: the structure is given by the main loop of the algorithm and is implementation dependent. We describe the PFA selection strategy of Z3-PFA in the 2nd paragraph of the sec 9.

6. In the evaluation, it is a bit unfortunate to see that there is no comparison with Trau [1-5]. This is critical. Lines 1039-1040 mention that it is because Trau does not ``support all string functions in our benchmarks, especially string-number conversion.'' This argument is unconvincing. In the evaluation, you can compare your approach against a vanilla Z3. And Trau is implemented on top of Z3 (see section 1 of [1]). Logically, Trau could be extended to handle the conversion (just by ``enabling'' some features in the vanilla Z3). If there is any principle limitation, it's worthwhile to justify it.

Response: Z3 is a DPLL(T) based solver, i.e., a DPLL engine + multiple theory solvers. The implementation of Trau is done by substituting the string solver of Z3 with a completely new one. Str-int conversion is not implemented in the new string solver of Trau, thus str-int conversion constraints will be internally discarded. If we run the str-int conversion benchmarks on Trau, we will either get exceptions or incorrect answers.

Reviewer C

1. I performed a very quick experiment in which I converted the function *checkLuhn* into C code, and find out KLEE works fine. This demonstrates that the state-of-the-art solvers are capable of handling the path constraints, which left me some negative feelings in the very beginning. Perhaps change the example program?

Response: Converting checkLuhn into C code will alter its semantics substantially.  The function makes use of array accesses with charAt() calls (lines 4, 8, 12) and, which, in JavaScript, require string-integer conversions to handle since the array state must be indexed with strings (see discussion starting at line 150 of the submission).  A C translation that did this would be very unnatural.

2. It would be better if the authors can provide more background to support your claim that string-number conversion is difficult for existing solvers. Is there an empirical study the authors can refer to? or educate the readers why those constraints are hard to solve.

Response: maybe we should answer the difficult and significancy of string-number conversion together at the beginning of the rebuttal.

3. I'm not clear about the "correct answer" (line 137), do you mean these solvers provide a wrong answer (e.g., return SAT for UNSAT constraints) or just return "unknown"? If they provide a wrong answer, why does it happen?

Response: We mean both. Sometime they just keep running without give any response. Sometimes, it provides a wrong answer. There might be several reasons that the answer is wrong. We checked the source code of Z3. One reason it gives wrong answer is that when both n and x in the constraint n=int(x) are variables, it over-approximates the int(x) function with an uninterpreted function. Sometimes Z3 gives a wrong model because of this over-approximation.

4. It is not clear to me how you construct a PFA that is aware of the domain of a constraint. For example, how do you infer that the domain of a constraint to be a binary number, octal number, decimal number or hexadecimal number? For a constraint "var[0] == var[1] && $\forall$ i in [0,len(var)-1], toNum(var[i]) > 9" that is only satisfiable for hexadecimal numbers, how do you know the domain is a hexadecimal number (i.e., [0, f]) as opposed to a decimal number (i.e., [0, 9])? Using PFA will simplify solving the string constraints. However, the performance of the under-approximation depends on the alphabet because the automaton consists of multiple one-way cycles (at line 1002). I can imagine the performance difference in a binary number, octal number, decimal number and hexadecimal number on which the size of each loop in PFA depends(e.g., the size of each loop is 8 for octal numbers). Therefore, would it be beneficial to convert different number systems to binary number systems?

Response: ...

5. At line 894, it is not clear to me how each numeric PFA covers a large and practically significant portion of numerals. In my view, each numeric PFA can only cover one number.

Response, in PFA, transition are labelled by variables, c.f., Fig 3. Each assignment of the variables corresponds to a number. E.g., assigning v_0...v_{m-1} to 0 and v_m to 1 corresponds to the integer number 1.

6. In the evaluation, you only give the run-time results for the evaluation on *checkLuhn* algorithm. This is rather strange, why don't you also provide solving time in table 1 or table 2? Additionally, why is its timeout budget (120s) different from the previous experiment (10s).

Response: We need to handle the time required for artifact evaluation. For table 1 and table 2, it has too many instances that an 120s timeout would require too much time to reproduce the same result. The evaluation of checkLuhn has only one instance so higher time budget is still possible.

7. I think "PASS: String Solving with Parameterized Array and Interval Automaton" is closely related, which I believe you should compare in your evaluation.

Some notes from reading: He might be right. The PASS uses quantified logic over arrays where the array length is a parameter (they have special quant. elim.), and they use automata which they encode to this kind of constraints, it resembles Presburger, and they do int-to-string in a way close to ours perhaps. It is also well cited, although nobody who cites them compares with them, and they do not compare much with anybody themselves, and the tool is nowhere to be found. 
We can perhaps say that their approach produces quantified formulae that are supposedly much harder to solve, and that our approach is much more flexible since it relates on SMT.

Response:

8. I am a bit struggling to learn the insight of your approach. In essence, why is your approach better than the existing solvers?

Response:

Reviewer D

1. - Although the solver is incomplete (which is not surprising for an undecidable problem in general), it would be useful to learn what heuristics were used in the iterative process of considering more expressive PFA. Was this process partially or fully automated? Do you prioritize among the PFAs of different variables?

Response: the PFA selection is an implementation. We describe the PFA selection strategy of Z3-PFA in sec 9, 2nd paragraph. The process is fully automated.

2. - It also seems that there may be some wasted effort when considering larger PFAs iteratively. Is some work re-usable? Could your solver be made incremental over the PFAs?

Response: we consider this as a future work.

3. - please explain requirement 3 that each character in \Sigma appears on at most one transition. Is it at most one transition in the full automata or in each loop?

Response: it is in the full automata